---
title: "Meta Data Report"
author: "Quanhu Sheng"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 3
    code_folding: hide
    number_sections: yes
---

```{r,eval=FALSE,echo=FALSE}
library(xfun)
root_dir = '/nobackup/h_cqs/shengq2/examples/20240411_meta_report'
dir.create(root_dir, showWarnings=FALSE)
setwd(root_dir)

file.copy('/nobackup/h_cqs/shengq2/program/ngsperl/lib/CQS/reportFunctions.R', './', overwrite=TRUE)
file.copy('/nobackup/h_cqs/shengq2/program/cqsperl/examples/20240411_meta_report.rmd', "./", overwrite=TRUE)

date_str = format(Sys.time(), "%Y%m%d")

if(1){
  xfun::Rscript_call(
    rmarkdown::render,
    list(input="20240411_meta_report.rmd",
        output_file=paste0(date_str, "_cohort_meta.html"))
  )
}
```

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
root_dir = "/nobackup/h_cqs/shengq2/examples/20240411_meta_report"

library(knitr)
library(Hmisc)
require(arsenal)
library(gtsummary)
library(flextable)
library(officer)

knitr::opts_chunk$set(
  echo=TRUE, 
  include=TRUE, 
  warning=FALSE, 
  message=FALSE, 
  results="asis"
)

date_str = format(Sys.time(), "%Y%m%d")

knitr::opts_knit$set(root.dir = root_dir)
setwd(root_dir)
```

```{r}
source('reportFunctions.R')

mono_meta_file = "/nobackup/h_cqs/shengq2/program/RaviProjects/20230526_mona_rnaseq/20240313_adiposity_mona_kb_rnaseq_for_obesity_revision.Mona_meta.csv"

kb_meta_file = "/nobackup/h_cqs/shengq2/program/RaviProjects/20230526_mona_rnaseq/20240313_adiposity_mona_kb_rnaseq_for_obesity_revision.KB_meta.csv"

PLHIV_meta_file_1 = "/nobackup/h_cqs/shengq2/program/RaviProjects/20230526_mona_rnaseq/HATIM_LassoScores.csv"
PLHIV_meta_file_2 = "/nobackup/h_cqs/shengq2/program/RaviProjects/20230526_mona_rnaseq/HATIM_metadata_5.8.23.csv"
```

# AHA cohort

Read meta file: `r basename(mono_meta_file)`

```{r}
aha_meta=data.frame(fread(mono_meta_file), row.names=1) |>
  dplyr::mutate(female_sex=sex=="Female",
         non_white=is.na(race) | race != "White") |> 
  dplyr::select(age, female_sex, non_white, bmi, homa.ir) |> 
  dplyr::rename(`Age (years)`=age,
    `Women`=female_sex,
    `Non-white race`=non_white,
    `Body mass index (kg/m2)`=bmi,
    `HOMA-IR`=homa.ir)
    
aha_meta |> tbl_summary() |> modify_header(all_stat_cols() ~ "Discovery cohort (N={N})")
```

There were `r nrow(aha_meta)` subjects in AHA meta file.

# VA cohort

Read meta file: `r basename(kb_meta_file)`

```{r}
va_meta = fread(kb_meta_file) |> 
  dplyr::mutate(
    female_sex=sex=="Female",
    non_white=race != 4) |> 
  dplyr::select(age, female_sex, non_white, bmi, homa.ir) |> 
  dplyr::rename(`Age (years)`=age,
    `Women`=female_sex,
    `Non-white race`=non_white,
    `Body mass index (kg/m2)`=bmi,
    `HOMA-IR`=homa.ir)
    
va_meta |> tbl_summary() |> modify_header(all_stat_cols() ~ "Validation cohort 1 (N={N})")
```

There were `r nrow(va_meta)` subjects in VA meta file.

# PLHIV cohort

Read meta file: `r basename(PLHIV_meta_file_1)` and `r basename(PLHIV_meta_file_2)`

```{r}
PLHIV_meta_1=fread(PLHIV_meta_file_1)
PLHIV_meta_2=fread(PLHIV_meta_file_2)
PLHIV_meta_2=PLHIV_meta_2[PLHIV_meta_2$hatim_clin_visit_pid %in% PLHIV_meta_1$HATIMID, c("hatim_clin_visit_pid", "meta_sex", "meta_age", "meta_bmi", "meta_race", "std_homa_ir")]
PLHIV_meta_2=data.frame(PLHIV_meta_2, row.names=1)

PLHIV_meta = PLHIV_meta_2 |> 
  dplyr::mutate(
    female_sex=meta_sex == 1,
    non_white=meta_race != 0) |> 
  dplyr::select(meta_age, female_sex, non_white, meta_bmi, std_homa_ir) |> 
  dplyr::rename(`Age (years)`=meta_age,
    `Women`=female_sex,
    `Non-white race`=non_white,
    `Body mass index (kg/m2)`=meta_bmi,
    `HOMA-IR`=std_homa_ir)

PLHIV_meta |> tbl_summary() |> modify_header(all_stat_cols() ~ "Validation cohort 2 (N={N})")
```

There were `r nrow(PLHIV_meta)` subjects in PLHIV meta file.

# Descriptive statistics

```{r}
all=rbind(aha_meta |> mutate(Cohort="AHA"), 
          va_meta |> mutate(Cohort="VA"), 
          PLHIV_meta |> mutate(Cohort="PLHIV"))
write.csv(all, "table1.clinical_summary.csv", row.names=FALSE)

all |> tbl_summary(by="Cohort") |> add_p() 
```

# Table 1

```{r}
add_stat_pairwise <- function(data, variable, by, ...) {
  # calculate pairwise p-values
  pw <- pairwise.t.test(data[[variable]], data[[by]], p.adj = "none")

  # convert p-values to list
  index <- 0L
  p.value.list <- list()
  for (i in seq_len(nrow(pw$p.value))) {
    for (j in seq_len(nrow(pw$p.value))) {
      index <- index + 1L
      
      p.value.list[[index]] <- 
        c(pw$p.value[i, j]) %>%
        setNames(glue::glue("**{colnames(pw$p.value)[j]} vs. {rownames(pw$p.value)[i]}**"))
    }
  }
  
  # convert list to data frame
  p.value.list %>% 
    unlist() %>%
    purrr::discard(is.na) %>%
    t() %>%
    as.data.frame() %>%
    # formatting/roundign p-values
    dplyr::mutate(dplyr::across(everything(), style_pvalue))
}

all_tbl = all |> 
  tbl_summary(by=Cohort) |> 
  add_stat(everything() ~ add_stat_pairwise) |>
  modify_header(stat_1="Discovery cohort\n(N=76)", 
                stat_2="Validation cohort 1\n(N=73)", 
                stat_3="Validation cohort 2\n(N=71)") |>
  as_flex_table()

table1_file = "table1.clinical_summary.docx"
all_tbl |>save_as_docx(path = table1_file, pr_section = prop_section(
    page_size = page_size(orient = "landscape")
  ))   
all_tbl
```

Table 1 was saved to [table1.clinical_summary.docx](table1.clinical_summary.docx).
