---
title: "Tabula Sapiens snRNA data"
author: 
- name: "Quanhu Sheng"
  affiliation: "CQS/Biostatistics, VUMC"
  email: "quanhu.sheng.1@vumc.org"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::readthedown:
    toc_depth: 3
    number_sections: yes
---

```{r,eval=FALSE,echo=FALSE}
library(xfun)
root_dir = '/data/cqs/references/Tabula_Sapiens/20240708_Tabula_Sapiens_snRNA_cellxgene'
dir.create(root_dir, showWarnings=FALSE)

setwd(root_dir)

file.copy('/nobackup/h_cqs/shengq2/program/ngsperl/lib/CQS/reportFunctions.R', "./", overwrite=TRUE)
file.copy('/nobackup/h_cqs/shengq2/program/ngsperl/lib/scRNA/scRNA_func.r', "./", overwrite=TRUE)

file.copy('/nobackup/h_cqs/shengq2/program/cqsperl/database/scRNA/20240708_Tabula_Sapiens_snRNA_cellxgene_sub.rmd', "./", overwrite=TRUE)
file.copy('/nobackup/h_cqs/shengq2/program/cqsperl/database/scRNA/20240708_Tabula_Sapiens_snRNA_cellxgene.rmd', "./", overwrite=TRUE)

date_str = format(Sys.time(), "%Y%m%d")

xfun::Rscript_call(
  rmarkdown::render,
  list(input="20240708_Tabula_Sapiens_snRNA_cellxgene.rmd",
       output_file=paste0(date_str, "_check_Tabula_Sapiens_snRNA_cellxgene.html"))
)
```

```{css, echo=FALSE}
#content{
    max-width:1920px;
}
```

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
root_dir = '/data/cqs/references/Tabula_Sapiens/20240701_check_Tabula_Sapiens_snRNA_cellxgene'

load_install<-function(library_name, library_sources=library_name){
  if(!require(library_name, character.only = T)){
    BiocManager::install(library_sources, ask=FALSE)
  }
  library(library_name, character.only = T)
}

load_install("knitr")
load_install("Hmisc")
load_install("arsenal")
load_install("gtsummary")
load_install("flextable")

knitr::opts_chunk$set(
  include=TRUE, 
  echo=FALSE, 
  warning=FALSE, 
  message=FALSE, 
  results="asis"
)

date_str = format(Sys.time(), "%Y%m%d")

knitr::opts_knit$set(root.dir = root_dir)
setwd(root_dir)

source('reportFunctions.R')
source('scRNA_func.r')
```

The Tabula Sapiens snRNA data is available at [cellxgene](https://cellxgene.cziscience.com/e/53d208b0-2cfd-4366-9866-c3c6114081bc.cxg/) and [rds file](https://datasets.cellxgene.cziscience.com/981bcf57-30cb-4a85-b905-e04373432fef.rds).

<pre>
https://cellxgene.cziscience.com/e/53d208b0-2cfd-4366-9866-c3c6114081bc.cxg/
https://datasets.cellxgene.cziscience.com/981bcf57-30cb-4a85-b905-e04373432fef.rds
</pre>

```{r}
obj_file="/data/cqs/references/Tabula_Sapiens/981bcf57-30cb-4a85-b905-e04373432fef.rds"
obj_10x3_file="/data/cqs/references/Tabula_Sapiens/cellxgene_10x3.rds"
```

# MD5 of files

```{r}
md5_tbl=rbind(
  data.frame("name"="obj_file", "file"=obj_file, "md5"="18113edf2ab892b2ccd8226cc10e7d8e")
)

for(i in 1:nrow(md5_tbl)){
  check_md5(md5_tbl$file[i], md5_tbl$md5[i])
}

print_table(md5_tbl, row.names=F)
```

# Data preparation

Load object from `r obj_file`.

```{r}
obj=readRDS(obj_file)
```

## Assay

```{r}
get_freq_table(obj@meta.data, "assay") |>
  print_table(row.names=F)
```

<mark>Since we will focus on gene expression in most of the data analysis using Tabula Sapiens data, 
we will use 10x 3' v3 assay only.</mark>

```{r}
obj = obj |> subset(assay == "10x 3' v3")
```

## Compartment

```{r}
get_freq_table(obj@meta.data, "compartment") |>
  print_table(row.names=F)
```

<mark>There are only 11 germ line cells in the object, we will remove those 11 germ line cells from the object.</mark>

```{r}
obj = obj |> subset(compartment != "germ line")
```

## Update meta

```{r}
obj@meta.data |>
  dplyr::filter(cell_type == "unknown") |>
  dplyr::select(cell_type, compartment) |>
  dplyr::count(cell_type, compartment) |>
  print_table(row.names=F)
```

<mark>Since all unknown cell types are in epithelial compartment, we will rename cell type of those cells to epithelial.</mark>

```{r}
obj@meta.data$cell_type=as.character(obj@meta.data$cell_type)
#obj@meta.data$cell_type[is.na(obj@meta.data$cell_type)] = "epithelial"
obj@meta.data$cell_type[obj@meta.data$cell_type == "unknown"] = "epithelial"
```

<mark>Rename tissue_in_publication to tissue, factorize donor_id and compartment, save the object to `r obj_10x3_file`.</mark>

```{r}
obj@meta.data$orig.tissue=obj@meta.data$tissue
obj@meta.data$tissue = factor_by_count(obj@meta.data$tissue_in_publication)
obj@meta.data$donor_id = factor_by_count(obj@meta.data$donor_id)
obj@meta.data$compartment = factor_by_count(obj@meta.data$compartment)
```

<mark>Replace gene id with gene feature name</mark>

```{r}
obj@assays$RNA@counts@Dimnames[[1]]=as.character(obj@assays$RNA@meta.features$feature_name)
obj@assays$RNA@data@Dimnames[[1]]=as.character(obj@assays$RNA@meta.features$feature_name)
obj$tissue=factor_by_count(obj$tissue)
```

<mark>Save the object to `r obj_10x3_file`.</mark>

```{r}
if(!file.exists(obj_10x3_file)){
  saveRDS(obj, obj_10x3_file)
}
```

## Donor

```{r}
get_freq_table(obj@meta.data, "donor_id") |>
  print_table(row.names=F)
```

<mark>You may notice that the top 2 donors contributed more than half cells.</mark>

# Meta data

## Summary

```{r}
meta = obj@meta.data |>
  dplyr::select(-observation_joinid, 
                -nFeaturess_RNA, 
                -nCounts_RNA_UMIs, 
                -assay_ontology_term_id, 
                -disease_ontology_term_id, 
                -is_primary_data, 
                -organism_ontology_term_id, 
                -suspension_type, 
                -cell_type_ontology_term_id, 
                -tissue_ontology_term_id, 
                -development_stage_ontology_term_id, 
                -self_reported_ethnicity_ontology_term_id, 
                -tissue_type, 
                -sex_ontology_term_id, 
                -manually_annotated)

s_tbl = meta |> tbl_summary() 

s_tbl |>
  as_gt() |> 
  gt::gtsave(filename = "meta.docx")

s_tbl
```

<mark>You can find the meta data summary in [meta.docx](meta.docx).</mark>

## Tissue ~ Compartment distribution

```{r}
table(obj$tissue, obj$compartment) |>
  as.data.frame.matrix() |>
  print_table(byDT=FALSE, row.names=TRUE)
```

# Compartment UMAP

```{r}
g<-get_dim_plot_labelby(obj, label.by="compartment", reduction="scvi_umap")
compartment_umap_file="compartment_umap.png"
ggsave(compartment_umap_file, g, width=10 + get_legend_width(g), height=10, units="in", dpi=300, limitsize=FALSE, bg="white")
include_graphics(compartment_umap_file)
```

# Tissue UMAP

## Global UMAP

```{r}
g<-get_dim_plot_labelby(obj, label.by="tissue", reduction="scvi_umap")
tissue_umap_file="tissue_umap.png"
ggsave(tissue_umap_file, g, width=10 + get_legend_width(g), height=10, units="in", dpi=300, limitsize=FALSE, bg="white")
include_graphics(tissue_umap_file)
```

Since the cells in UMAP were clustered together by cell type rather than tissue, it's really hard to say where the tissue is located in the UMAP plot.

```{r}
tissues=levels(obj@meta.data$tissue)

umap_content=""
for(tissue in tissues){
  umap_content = paste0(umap_content, "```{r}\ntissue='", tissue, "'\n```\n\n```{r child='20240708_Tabula_Sapiens_snRNA_cellxgene_sub.rmd'}\n```\n\n")
}
umap_rmd="tissue_umap.rmd"
writeLines(umap_content, umap_rmd)
```

```{r child=umap_rmd}
```

# Final object MD5

```{r}
md5_tbl=rbind(
  data.frame("name"="obj_10x3_file", "file"=obj_10x3_file, "md5"="")
)

for(i in 1:nrow(md5_tbl)){
  md5_tbl$md5[i] = check_md5(md5_tbl$file[i], "", return_md5=TRUE)
}
print_table(md5_tbl, row.names=F)
```

# Save the session information

```{r}
writeLines(capture.output(sessionInfo()), paste0(date_str, '.sessionInfo.txt'))
```
